Review and comparison of tools for managing X.509 certificates and keys

### 2016-12-01 TEX/draft.tex
Changed everything what we talked about today

\title{Review and comparison of tools for managing X.509 type keys and certificates}
\author{Oliver Bodnár}
\date{November 2016}

@@

\begin{itemize}
\item \textit{Yes} - The tool supports signing of certificate signing request.
\item \textit{No} - the tool does not support signing of certificate signing requests.
\end{itemize}

@@

\begin{itemize}
 \item \textit{Yes} - if it is possible to combine private key and corresponding chain into a single file
 \item \textit{No} - otherwise.
\end{itemize}

@@

\begin{itemize}
 \item \textit{Yes} - Possibility of choosing how long the certificate will be valid. Done by certificate authority.
 \item \textit{No} - If there is no option to change how long the certificate will be valid, so the duration is a fixed ammount.
\end{itemize}

@@

      \paragraph{Specify Type}
To specify if generated certificate belongs to CA or is end certificate you choose \verb+<BOOL_VAL>+ true or false respectively.
      \paragraph{Specify path length}
Specifying of path length is made by setting a \verb+<maxChainLengthInteger>+ to desired value (e.g. if generated certificate/key should be used only for signingend certificates the value is 0)
  \subsection{Specifications}

@@

    To export certificate chain from a PKCS\#12 file use:
\begin{verbatim}
openssl pkcs12 -in name.[pfx|p12] -nokeys -cacerts -out CAchain.pem
\end{verbatim}
    It should be mentioned that if there are multiple certificates in a chain they are all going to be in same output file.

@@

@@

      \paragraph{Certificate Signing Request}
Possible by using \verb+gpgsm --gen-key > <output_file>+ command, but private key is not a part of output CSR.

GnuPG is unable to do anything besides creating a certificate signing request in X.509 public key infrastructure

@@

  \subsection{Type}
Keystore Explorer supports both Java Keystore and PKCS\#12 type files, in addition it has a support for alternatives to JKS such as JCEKS, BKS, BKS-V1, UBER.
\begin{figure}[!ht]
 \caption{Choosing type}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/Type.jpg}
\end{figure}

@@

\begin{figure}[!ht]
 \caption{Generating keys}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/genKeys.jpg}
\end{figure}

\begin{figure}[!ht]
 \caption{Generating certificate}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/genCert.jpg}
\end{figure}

@@

\begin{figure}[!ht]
 \caption{Generating a CSR}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/genCSR.jpg}
\end{figure}

@@

\begin{figure}[!ht]
 \caption{Basic Constraints}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/Basic_Constraints.jpg}
\end{figure}

@@

\begin{figure}[!ht]
 \caption{Signing of a CSR}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/CSRsign.jpg}
\end{figure}

@@

\begin{figure}[!ht]
 \caption{Generation of key for new cert}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/genKey.jpg}
\end{figure}

\begin{figure}[!ht]
 \caption{Specify usage}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/genKey_specif.jpg}
\end{figure}

      \paragraph{Certificate Signing Request}
Certificate Signing Request is generated similarly as a certificate itself, only difference is that you go under Certificate signing request tab and press New Request.
\begin{figure}[!ht]
 \caption{Generation of CSR}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/newCSR.jpg}
\end{figure}

      \paragraph{Specify length}
Length of keys is specified during generation of keys.
      \paragraph{Specify algorithm}
Algorithm used for key generation is specified during generation of keys (\verb+RSA|DSA|EC+).
\begin{figure}[!ht]
 \caption{Specifying of length and algorithm used}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/genKeylength.jpg}
\end{figure}

@@

\begin{figure}[!ht]
 \caption{Basic constraints, validity and SAN}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/BC_ext.png}
\end{figure}

@@

    \subsubsection{Create combination of private key and signed chain}
If the private key is provided PEM encoded it can be imported to XCA's database. If not you need to generate new key pair, then use export option to export private part of a key, delete key pair from database and then import private key. to create a combination, simply generate new certificate while using said private key.
    \subsubsection{Specify certificate validity}
Time period during which a certificate is valid is defined under Extensions tab.
    \subsubsection{Setting Subject Alternative Name for end certificates}
Subject Alternative Names are specified in appropriate section under an Extensions tab.
\textit{(Figure 6.5)}

@@

    \subsubsection{Private key only}
Private key is stored and generated while generating a certificate but the type .pvk is uncommon.

@@

\section{OpenSSL}
It is a great command line tool, can work with PEM, DER and PKCS\#12 files. Strong point is an easy form of creating a self-signed certificate, however using it for CA certificates with ca option is not ideal -  creating CA is not advised in O'REILLY's Network Security with OpenSSL by J. Viega, M. Messier and P. Chandra on page 59,
\begin{displayquote}
\textit{Since OpenSSL's command-line CA functionality was intended primarely as an example of how to use OpenSSL to build a CA, we don't recommend that you attempt to use it in a large production environment.}
\end{displayquote}
and also hinted in manual pages:
\begin{displayquote}
\textit{The ca command is quirky and at times downright unfriendly.}

>>>>

\title{Review and comparison of tools for managing X.509 certificates and keys}
\author{Oliver Bodnár}
\date{November 2016}

@@

\textit{Yes} means that the tool supports signing of certificate signing request. \textit{No} if it does not support signing of certificate signing requests.

@@

\textit{Yes} if it is possible to combine private key and corresponding chain into a single file, \textit{No} - otherwise.

@@

\textit{Yes} - Possibility of choosing how long the certificate will be valid. Done by certificate authority. \textit{No} - If there is no option to change how long the certificate will be valid, so the duration is a fixed ammount.

@@

      \paragraph{Specify Type}
To specify if generated certificate belongs to CA or is end certificate you choose \verb+<BOOL_VAL>+ true or false respectively.
      \paragraph{Specify path length}
Specifying of path length is made by setting a \verb+<maxChainLengthInteger>+ to desired value (e.g. if generated certificate/key should be used only for signing end certificates the value is 0)
  \subsection{Specifications}

@@

    To export certificate chain from a PKCS\#12 file use:
\begin{verbatim}
openssl pkcs12 -in name.[pfx|p12] -nokeys -cacerts -out CAchain.pem
\end{verbatim}
    It should be noted that if there are multiple certificates in a chain they are all going to be in same output file.

@@

\section{Note}
However using self-signed certificate made with \verb+ca+ option as CA certificates is not ideal -  creating CA is not advised in O'REILLY's Network Security with OpenSSL by J. Viega, M. Messier and P. Chandra on page 59,
\begin{displayquote}
\textit{Since OpenSSL's command-line CA functionality was intended primarely as an example of how to use OpenSSL to build a CA, we don't recommend that you attempt to use it in a large production environment.}
\end{displayquote}
and also hinted in manual pages:
\begin{displayquote}
\textit{The ca command is quirky and at times downright unfriendly.}

@@

      \paragraph{Certificate Signing Request}
Possible by using \verb+gpgsm --gen-key > <output_file>+ command, but private key is not a part of output CSR.     
\section{Note}
GnuPG is unable to do anything besides creating a certificate signing request in X.509 public key infrastructure, that does not include a private key within itself.

@@

  \subsection{Type}
Keystore Explorer supports both Java Keystore and PKCS\#12 type files, in addition it has a support for alternatives to JKS such as JCEKS, BKS, BKS-V1, UBER.
\begin{figure}[!htbp]
 \caption{Choosing type}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/Type.jpg}
\end{figure}

@@

\begin{figure}[!htbp]
 \caption{Generating keys}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/genKeys.jpg}
\end{figure}

\begin{figure}[!htbp]
 \caption{Generating certificate}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/genCert.jpg}
\end{figure}

@@

\begin{figure}[!htbp]
 \caption{Generating a CSR}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/genCSR.jpg}
\end{figure}

@@

\begin{figure}[!htbp]
 \caption{Basic Constraints}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/Basic_Constraints.jpg}
\end{figure}

@@

\begin{figure}[!htbp]
 \caption{Signing of a CSR}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/Keystore_Explorer/CSRsign.jpg}
\end{figure}

@@

\begin{figure}[!htbp]
 \caption{Generation of key for new cert}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/genKey.jpg}
\end{figure}

\begin{figure}[!htbp]
 \caption{Specify usage}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/genKey_specif.jpg}
\end{figure}

      \paragraph{Certificate Signing Request}
Certificate Signing Request is generated similarly as a certificate itself, only difference is that you go under Certificate signing request tab and press New Request.
\begin{figure}[!htbp]
 \caption{Generation of CSR}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/newCSR.jpg}
\end{figure}

      \paragraph{Specify length}
Length of keys is specified during generation of keys.
      \paragraph{Specify algorithm}
Algorithm used for key generation is specified during generation of keys (\verb+RSA|DSA|EC+).
\begin{figure}[!htbp]
 \caption{Specifying of length and algorithm used}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/genKeylength.jpg}
\end{figure}

@@

\begin{figure}[!htbp]
 \caption{Basic constraints, validity and SAN}
 \centering
  \includegraphics[width=0.5\textwidth]{../Dependancies/XCA/BC_ext.png}
\end{figure}

@@

    \subsubsection{Create combination of private key and signed chain}
If the private key is provided PEM encoded it can be imported to XCA's database. If not you need to generate new key pair, then use export option to export private part of a key, delete key pair from database and then import private key. To create a combination, simply generate new certificate while using said private key.

To export certificate with private key and certificate chain, simply use Export option under \textit{Certificates} tab. There you can choose if you wish to export it in a single \textit{PKCS\#12} type file or certificate and key separately, PEM encoded.

    \subsubsection{Specify certificate validity}
Time period during which a certificate is valid is defined under Extensions tab.
    \subsubsection{Setting Subject Alternative Name for end certificates}
Subject Alternative Names are specified in appropriate section under an Extensions tab.
\textit{(Figure 6.5)}

@@

    \subsubsection{Private key only}
Private key is stored and generated while generating a certificate but the type .pvk is uncommon. It is Microsoft's own key storage type more on \url{http://www.drh-consultancy.demon.co.uk/pvk.html}

@@

\section{OpenSSL}
It is a great command line tool, can work with PEM, DER and PKCS\#12 files. Strong point is a easy form of
creating a self-signed certificate, however using it for CA certificates with ca option is not ideal - it is noted in man
pages and also a O’REILLY OpenSSL book.
